---
title: "Data Visialization"
author: "Anna Nesterovich"
date: '1February 14th, 2019 г '
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reformat data for plotting 

Alaska Department of Fish and Game, Division of Commercial Fisheries. 2017. Daily salmon escapement counts from the OceanAK database, Alaska, 1921-2017. Knowledge Network for Biocomplexity. doi:10.5063/F1S46Q6M.

```{r}
data_url <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e"

#looks for the file on disk first, then downloads
esc <- tryCatch(
    read.csv("data/escapement.csv", stringsAsFactors = FALSE),
    error=function(cond) {
        message(paste("Escapement file does not seem to exist, so get it from the KNB."))
        esc <- read.csv(url(data_url, method = "libcurl"), stringsAsFactors = FALSE)
        return(esc)
    }
)

head(esc)
```

```{r}
esc <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", method="libcurl"), stringsAsFactors=F)
head(esc)
```

All the libraries we'll need 
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(leaflet)
library(DT)
```


* calculate annual escapment by species and region
* filter for just Pacific salmon (Sockeye, Chinook, Coho, Pink, Chum)

annual_esc
use functions:
gather
spread
rename
mutate
select
filter
group_by
summarize
arrange
separate
unite
split-apply-combine (region as grouping variable)
distinct - pulls out unique row or combinations of rows

```{r}
#separate for the date to extract year from it
#group_by - we want to group by region and year and species
#summarize - to sum all the dily counts into annual counts for a species
    #in a region for a year
#filter - pull out the species we want (the 5 Pacific ones)
#Species %in% c(...), not the or| thing would be better

annual_esc <- esc %>% 
  separate(sampleDate, c("year", "month", "day"), "-", remove=F) %>% 
    #if I don't want to ungroup later, I need to do year as numeric here
    group_by(SASAP.Region, Species, year) %>% 
    summarize(annualEsc = sum(DailyCount)) %>% 
    filter(Species == "Sockeye" | Species == "Chinook" |
               Species =="Coho" | Species =="Pink" | Species=="Chum") %>% 
    #need to coerce year to numeric type
    ungroup() %>% 
    mutate(year=as.numeric(year))
annual_esc

```

So now the data is tidy enough and we are going to make some plots

# Generate some static plots
first is the ggplot call with aes stand for aestetics, what you want to add to the plot and how it looks like
the second part is geometry, to tell it how to plot your data there, some sort of geom function and there are a lot, we'll do bar plot
color is the outline, fill the color within, color goes not in the aes, not a variable in the dataset, it goes to geom
```{r}
ggplot(annual_esc, aes(x = Species, y = annualEsc)) +
    geom_col(color="blue", fill="cornflowerblue")
```

Can put almost everything into geom, when you say want different geometries on the the same plot (bar and a line say), or two diff datasets on the same plot
```{r}
ggplot(annual_esc) +
    geom_col(mapping=aes(x = Species, y = annualEsc), color="blue", fill="cornflowerblue")
```
scale_fill_manual can specify colors in pairs
there is a good cheatsheet for ggplot vizulation in help
```{r}
ggplot(annual_esc, aes(x = Species, y = annualEsc, fill=SASAP.Region)) +
    scale_fill_brewer(palette="Dark2") +
    geom_col()
```

```{r}
#I can save all the theme settings and use it for a bunch of plots
my_theme <- theme_bw() +   #theme_test looks very good
    theme(legend.position="bottom", legend.title = element_blank(),
          plot.title=element_text(hjust=0.5, color="green"),
              axis.text.y = element_text(angle = 45))
```
First theme_test or whatever, then theme. Otherwise the theme will be overridden. Cll theme only once

```{r}
#just Kodiak
ggplot(filter(annual_esc, SASAP.Region=="Kodiak"), aes(x=year, y=annualEsc, color=Species)) +
    geom_line() +
    geom_point() +
    ylab("Escapment (num of fish") +
    xlab("Year") +
    ggtitle("Kodiak Salmon Escapment") +
    scale_y_continuous(labels=comma) +  #comma, percent
    my_theme
```
If when reading the file, there wasn't stringsAsFactors=F, there would be a problem converting to numbers. Or maybe not.


```{r, fig.width=5, fig.height=8}
#that sets the size of the picture I want in inches
p <- ggplot(annual_esc, aes(x=year, y=annualEsc, color=Species)) +
    geom_line()+
    geom_point() +
    scale_y_continuous(label=comma)+
    facet_wrap(~SASAP.Region, scales="free_y", ncol=2) +
    my_theme
p

#SAVING PLOTS

ggsave("Figures/regional_escapment.tiff", plot=p, width=7, height=8, units="in", dpi=400)

#will not ask overwriting the graph with a new code
```






# Generate an interactive map

Map the sampling locations for escapement

8 find (create a dataframe) distinct lat/long/location combinations

```{r}
locations <- esc %>% 
    distinct(Location, Latitude, Longitude) %>% 
    drop_na()

head(locations)
```

```{r, echo=F}
datatable(locations)
```

Let's make a map!


```{r}
leaflet(locations) %>% 
    addTiles() %>%  #just draws the map !!!!! Remember the projection (markater)
    addMarkers(lng= ~Longitude, lat = ~ Latitude, popup = ~ Location)
```



How to change spacing in RMarckdown
? html break tag
<br/>

leaflet is not build to generate static maps, use other packs
such as gglot
geom_sf for vector shapefiles
geom_raster



How to change the basemap tiles? Couldn't figure out right away. Too many options in the package -> not a very good documentation.


Change markers from that default droplets. including custom (like tiny diatoms?) 
