---
title: "TidySalmon"
author: "Anna Nesterovich"
date: '13 февраля 2019 г '
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyr)

#to call stats filter - stats::filter()
#to call stats lag - stats::lag()
#also masked from the base: intersect, setdiff, setequal, union
```


# Reformat Catch Data

* remove "all" column
* create species column
    * move from wide to long format
*general QA

Grabbing the data from
[Mike Byerly. Alaska commercial salmon catches by management region (1886- 1997). Gulf of Alaska Data Portal. df35b.304.2.](https://knb.ecoinformatics.org/view/df35b.304.2)

```{r eval=FALSE}
catch_original <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")
```

If that doesn't work, this should
```{r}
catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                    stringsAsFactors = FALSE)
head(catch_original)
```

Remove the 'all' and 'notesRegCode' columns using 'select'
Remember about the pipe operator %>%

So df_filtered <- filter(df, ...)
df_selected <- select(df_filtered, ...)

translates into 

df_cleaned <- df %>% 
              filter(...) %>%
              select(...)
              
Which can be very useful in my Shiny App.

To print %>% - ctrl + shift + m

```{r}
catch_long <- catch_original %>% 
    select(-All, -notesRegCode) %>%  #select comuns you want moved from wide to long
    gather(key="species", value="catch", -Region, -Year) #look at all columns escept Region and year
    

# it could be select select(Region, Year, and so on, but if there more,
#it's easier to 'de-select')

head(catch_long)
```

```{r}
#want to move back to the wide format
catch_wide <- catch_long %>% 
    spread(key=species, value = catch)
head(catch_wide)
```

```{r}
#or we can do
catch_wide <- catch_long %>% 
    spread(key="Region", value = "catch")
head(catch_wide)
```

## Clean upour data
* rename catch to catch_thousands
* change catch column to numeric
* create a new catch column in units num. (`#`) of fish

```{r, eval=F}
catch_clean <- catch_long %>% 
    rename(catch_thousands = catch) %>% #new name = old name
    mutate(catch_thousands = as.numeric(catch_thousands)) %>%  #couldn't do all - there were letters
    #mutate(catch = catch_thousand * 100)
head(catch_clean)
```

Testing chunk to figure out why turning numeric failed
```{r, eval=F}
test <- as.numeric(catch_long$catch)
i <- which(is.na(test) == T)
i
catch_long[i,]
mutate(catch_thousands = ifelse(catch_thousands == "I", 1, catch_thousands))
```

```{r}
catch_clean <- catch_long %>% 
    rename(catch_thousands = catch) %>%
    mutate(catch_thousands=ifelse(catch_thousands %in% c("I","M","Z"), 1, catch_thousands)) %>% 
    mutate(catch_thousands = as.numeric(catch_thousands)) %>%
    mutate(catch = catch_thousands * 100) %>% 
    select(-catch_thousands)
head(catch_clean)
```

## Split - Apply - Combine

* calculate mean catch by species
```{r}
species_mean <- catch_clean %>% 
    group_by(species, Region) %>% 
    summarise(catch_mean = mean(catch),
              num_obs= n())

head(species_mean)
```

```{r}
species_mean <- catch_clean %>% 
    group_by(species) %>% 
    summarise(catch_mean = mean(catch),
              num_obs= length(which(catch != 0)))

head(species_mean)
```

```{r}
Year_mean <- catch_clean %>% 
    group_by(Year, species) %>% 
    summarise(catch_mean = mean(catch)) %>% 
    arrange(-Year, desc(species)) %>% 
    filter(Year >= 1990 & Year <= 1996 )  ## &- and; | - or
Year_mean
```


# Join the Region Definitions

# Misc. Functions